% !TEX encoding = UTF-8 Unicode
%!TEX root = thesis.tex
% !TEX spellcheck = en-US
%%=========================================
\addcontentsline{toc}{section}{Executive Summary}
\section*{Executive Summary}
% General Introduction Paragraph 
The commercial Internet has radically changed the way we live and communicate over the last decade. The technology advances are in large part a positive contribution to our society, and stands as a fundamental building block for further advancement and progression in all aspects of human development. It is now also the primary information source for most humans, and being a free platform for opinions and discussion, the web is a strong force to change the publics opinion through manipulated and fabricated data. 

% General Paragraph on Social Media Bot Manipulation
Social media has become a vital part of our everyday life, and there is great value in generating followers, likes and engagement to increase the reach of your channel or message. This has created something called social bots, programs that automatically take actions on social media platforms to simulate broad support or hide dissatisfaction for an entity, opinion or product.

% Summary based on Emil's contribution - Bot Metrics
Many online customers make choices based on recommendations and reviews alone, and are therefor vulnerable to metrics manipulated by bots when shopping online. In addition, research shows that around 400,000 bots where actively influencing the 2016 US election on Twitter, giving bots not only economical benefit but possibly political influence. Research also tells us that people perceive some bots very trustworthy, reinforcing the believe that bots indeed successfully influence us through their actions.  

% Summary based on Jan's contribution
Both personal and corporate computer systems are targets of botnet recruiting, but there is several precautions to be taken to reduce the risk of becoming a botnet agent. For organizations, the security measures to be taken is dependent on many factors  like size, technology and resources available. We recommend a cooperation between the organization and available governmental organizations like CERT' and law enforcement, local providers like ISP's, and international organizations like Europol.

% Summary based on Fabian's contribution - Bot detection
Different approaches can be used to detect social bots, and they can be classified as Social Network Based, Crowd-Sourcing Based and Feature Based. Social Network Based approaches model the relationships between all users, and use the relationships to detect bot users. Crowd-Sourcing on the other hand, uses the social competence of actual humans to decide whether a given user is a bot, and Feature Based approaches use machine learning to detect bot users. All the approaches have its own weaknesses, and a combination is advisable to increase detection rates. It's also important to continue research since the number of bots is expected to increase and bot creators will continue to develop more sophisticated methods to make their bots work undetected.
\\ \\
% Summary based on Audun's contribution - Social Media Attacks on Computers vs Mobile
According to some sources, social media is used in large part from mobile platforms. Reports claim that 79\% of social media usage is on mobile platforms, and still rapidly gaining terrain. In present days it is much safer to use social media on a regular computer rather than on a mobile device, but its still depends to a large degree on user behavior and precaution. The easiest and most commonplace solution for companies today is to implement a Mobile Device Management System. This will separate the personal content from the work related content, and keep the work related content encrypted. In addition this gives the possibility to remotely erase data when devices are lost or stolen.

% General Paragraph on attacks
The Internet was originally built for communication between military entities, and was secured trough physical security like walls and locked doors. As the Internet has grown to be used by nearly everyone and connected to increasingly many things, security has become a major concern. In addition, the web stack has evolved to serve feature rich and dynamic experiences within the browser by expanding on the legacy applications, creating more challenges in the security aspects of the world wide web.

% Summary based on Jorgen's contribution - Common attack surfaces for web applications
The company's web application is often an important part of the business, and breaches could leak sensitive data or infect customers with malware through exploit kits. The most common vulnerabilities today are Cross-Site Scripting, SQL Injection and vulnerable JavaScript Libraries, and the vulnerabilities are easy to avoid by having a good focus on developer awareness, and possibly using vulnerability scanners and services from web security companies. It is important to to focus on a proactive approach to web application security, and the recommendation is to educate all developers on web security and if the size and complexity of the web application requires it, invest in a vulnerability scanner service.
        